### 2020-06-04 writeup/transfer.tex
sec 3

%We provide precise explanations to the phenomena of negative transfer in multi-task learning.
We provide tight bounds on the bias and variance of the multi-task estimator for two tasks.
We show theoretical implications for understanding the performance of multi-task learning.
(a) \textit{Task similarity}: we explain the phenomenon of negative transfer precisely as task models become different.
(b) \textit{Sample size}: we further explain a curious phenomenon where increasing the source sample size helps initially, but hurts eventually.
(c) \textit{Covariate shift}: as source sample size increases, we show that covariate shift worsens the performance of the multi-task estimator.
%We explain from three perspectives, including \textit{task similarity}, \textit{sample size} and \textit{covariate shift}.
%We show how negative transfer occurs by varying task similarity or sample size.
%Then we show that when source task sample size becomes large, covariate shift causes more negative effects.

@@

\begin{lemma}\label{lem_cov_shift_informal}

@@

\begin{theorem}\label{thm_main_informal}

@@

>>>>

%We provide precise explanations to the phenomena of negative transfer in multi-task learning.
We provide tight bounds on the bias and variance of the multi-task estimator for two tasks.
We show theoretical implications for understanding the performance of multi-task learning.
(a) \textit{Task similarity}: we explain the phenomenon of negative transfer precisely as task models become different.
(b) \textit{Sample size}: we further explain a curious phenomenon where increasing the source sample size helps initially, but hurts eventually.
(c) \textit{Covariate shift}: as source sample size increases, we show that covariate shift worsens the performance of the multi-task estimator.
Finally, we extend our result from two tasks to many tasks with the same features.
%We explain from three perspectives, including \textit{task similarity}, \textit{sample size} and \textit{covariate shift}.
%We show how negative transfer occurs by varying task similarity or sample size.
%Then we show that when source task sample size becomes large, covariate shift causes more negative effects.

@@

\begin{lemma}[Variance bound]\label{lem_cov_shift_informal}

@@

\begin{theorem}[Two tasks]\label{thm_main_informal}

@@

\subsection{Extensions}

Next, we describe our result for more than two tasks with same features, i.e. $X_i = X$ for any $i$.
This setting is prevalent in applications of multi-task learning to image classification, where there are multiple prediction labels/tasks for every image \cite{chexnet17,EA20}.
\begin{theorem}[Many tasks]\label{thm_many_tasks}
%Suppose $X=Z\Sigma^{1/2}\in \R^{n\times p}$ satisfy Assumption \ref{assm_secA1} with $\rho:=n/p>1$ being some fixed constant. Consider data models  $Y_i = X\beta_i + \varepsilon_i$, $i=1,2,\cdots, t$, where $\e_i\in \R^{n}$ are random vectors with i.i.d. entries with mean zero, variance $\sigma^2$ and all moments as in \eqref{assmAhigh}. Moreover, assume that $X$, $\beta_i$ and $\e_i$ are all independent of each other.
	%Let $n = c \cdot p$.
	%Let $X\in\real^{n\times p}$ and $Y_i = X\beta_i + \varepsilon_i$, for $i = 1,\dots,k$.
%	Consider $t$ data models $Y_i = X\beta_i + \varepsilon_i$, $i=1,2,\cdots, t$, where $X$ has covariance matrix $\Sigma$, and the entries of $\e_i$ are i.i.d. with mean zero and variance $\sigma^2$.
	%that satisfy Assumption \ref{assm_secA2} in the appendix.
	For the setting of $t$ tasks where $X_i = X$, for all $1\le i\le t$.
	Let $B^\star := [{\beta}_1,{\beta}_2,\dots,{\beta}_{t}]$ and $U_r\in\real^{t\times r}$ denote the linear model parameters.
	Let $U_r U_r^{\top}$ denote the best rank-$r$ subspace approximation of $(B^{\star})^\top\Sigma B^{\star}$.
	Assume that $\lambda_{\min}({B^{\star}}^\top\Sigma B^{\star})\gtrsim \sigma^2$.
	Let $v_i$ denote the $i$-th row vector of $U_r$.
	We have the following
	\squishlist
		\item  If	$\left(1 - \norm{v_t}^2 \right)\frac{\sigma^2}{\rho - 1} > \norm{\Sigma (B^{\star} U_r v_t - \beta_t)}^2 +  \oo \left( \|B^\star\|^2 + \sigma^2\right)$, then w.h.p $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$
		\item If $\left(1 - \norm{v_t}^2\right)\frac{\sigma^2}{\rho - 1} < \norm{\Sigma(B^{\star} U_r v_t - \beta_t)}^2 -  \oo \left( \|B^\star\|^2 + \sigma^2\right)$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\squishend
\end{theorem}
Theorem \ref{thm_many_tasks} shows that when all tasks share the same features but have different labels, we can derive a sharp transition from positive to negative transfer.
Specially, $(1 - \norm{v_t}^2)\sigma^2/(\rho - 1)$ shows the amount of variance reduced using the multi-task estimator.
On the other hand, $\norm{\Sigma (B^{\star} U_r v_t - \beta_t)}$ shows the bias of the multi-task estimator.
The proof of \ref{thm_many_tasks} can be found in Appendix \ref{app_proof_many_tasks}.
